Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-03-22 WSSSPE2_report/WSSSPE2_review.tex
undoing a few edits, most are good, will add requested cites later

\subsubsection{First breakout discussion: Group A}
The first sub-group discussed about the current system for credit and current
incentives, which it called ``hacking the incentive structure.'' The group
considered four potential points of leverage:

@@

After this discussion, group A brainstormed about incentives,
with the following items suggested:
\begin{itemize}
\item running programming contests, creating bounties for contributing to open
source software, etc.
\item augment author lists to give credit to people who do not now get credit
(and making them machine readable)
\item developing a microcitation standard and mechanism (for both software and
data)
\item developing a well-defined standard for author contribution -- what level
of contribution rises to the level of authorship?
\item leveraging social media for citation and reviewing of content -- then
using social media to bring more people into the review process than is
traditional
\item determining where else software can be cited and recorded (e.g.,
acknowledgments sections of papers)
\item developing a taxonomy of contributors, tied to places that these metrics
are already stored (e.g., ORCID) \choinote{cite}
\item making metadata easier to add for software, creating incentive for
providing software metadata -- note that this cannot be centralized
\item creating something like the h-index that tenure committees can make use of
- simplify a way of measuring and documenting the overall credit given to an
individual over different projects
\item thinking about publishing software versus journal - software does not have to
be novel
\item determining guidelines for recommending software characteristics for
tenure - perhaps draft guidelines then get ACM or IEEE agreement
\end{itemize}

@@

The discussion of incentivizing those in science acknowledged that the
publications system was far from perfect for incentivizing good software work.
Nonetheless discussion focused on bringing software people into publications.
There were two main suggestions. The first is to focus on end users of software
and encourage them to cite software the ``right" way. James Howison suggested
that his research showed that few projects were making a formal request for
citation (but that authors were not necessarily following those suggestions
anyway) \cite{howison2015jasist}. He suggested making access to the software
conditional on a license that requires citation. Others found this ``too
confrontational" and preferred to concentrate on making it easier to do the
right thing. The second was focusing on those leading software projects, and the
group was more enthusiastic about ``forcing" PIs to include their ``software
people" on publications, although there were few ideas on how exactly to do
this. Another technique mentioned was that when scientific software projects are
hosted in organizations like Apache, the scientific contributors can benefit
from building their reputations, perhaps yielding job offers that they can use
to negotiate better job and career packages.

@@

The group came up with the following actions to be performed:
\begin{enumerate}
\item Build a tool that can automatically determine who the contributors are 
(from a git or other repo), then allows the user to manually edit the output to 
add/remove people, define roles. 
\item Work with repositories to encourage them to provide the information we 
need based on what they already store.
\item Define what a citation file should look like and what it should be called.
\item Test adoption, for example, with LBNL.
\item Create credit file for a set of software.
\item Build a validator (and perhaps a visualizer) for credit files.
\item Write a tool to collect files and visualize/output interconnections (which 
software is used with which), based on an existing project.
\item When we (the group members) write papers, we should track the software 
we use, and encourage the software developers to make their software citable 
and create credit files.
\item Build a tool to export the credit file to BibTex and other citation styles.
\item Make sure the BibTex entries (exported from internal data) are somewhat 
standardized so that they can be imported into papers.  Also make sure that 
standard \LaTeX style files understand and accept these entries.
\end{enumerate}

>>>>

\subsubsection{First breakout discussion: Group A}
The first sub-group discussed issues around the current system for credit and current
incentives, which it called ``hacking the incentive structure.'' The group
considered four potential points of leverage:

@@

After this discussion, group A brainstormed about incentives,
with the following items suggested:
\begin{itemize}
\item running programming contests, creating bounties for contributing to open
source software, etc.
\item augment author lists to give credit to people who don't now get credit
(and making them machine readable)
\item developing a microcitation standard and mechanism (for both software and
data)
\item developing a well-defined standard for author contribution -- what level
of contribution rises to the level of authorship?
\item leveraging social media for citation and reviewing of content -- then
using social media to bring more people into the review process than is
traditional
\item determining where else software can be cited and recorded (e.g.,
acknowledgments sections of papers)
\item developing a taxonomy of contributors, tied to places that these metrics
are already stored (e.g., ORCID) \choinote{cite}
\item making metadata easier to add for software, creating incentive for
providing software metadata -- note that this cannot be centralized
\item creating something like the h-index that tenure committees can make use of
- simplify a way of measuring and documenting the overall credit given to an
individual over different projects
\item thinking about publishing software versus journal - software does not have to
be novel
\item determining guidelines for recommending software characteristics for
tenure - perhaps draft guidelines then get ACM or IEEE agreement
\end{itemize}

@@

The discussion of incentivizing those in science acknowledged that the
publications system was far from perfect for incentivizing good software work.
Nonetheless discussion focused on bringing software people into publications.
There were two main suggestions. The first is to focus on end users of software
and encourage them to cite software the ``right" way. James Howison suggested
that his research showed that few projects were making a formal request for
citation (but that authors weren't necessarily following those suggestions
anyway) \cite{howison2015jasist}. He suggested making access to the software
conditional on a license that requires citation. Others found this ``too
confrontational" and preferred to concentrate on making it easier to do the
right thing. The second was focusing on those leading software projects, and the
group was more enthusiastic about ``forcing" PIs to include their ``software
people" on publications, although there were few ideas on how exactly to do
this. Another technique mentioned was that when scientific software projects are
hosted in organizations like Apache, the scientific contributors can benefit
from building their reputations, perhaps yielding job offers that they can use
to negotiate better job and career packages.

@@

The group came up with the following actions to be performed:
\begin{enumerate}
\item Build a tool that can automatically determine who the contributors are 
(from a git or other repo), then allows the user to manually edit the output to 
add/remove people, define roles. 
\item Work with repositories to encourage them to provide the information we 
need based on what they already store.
\item Define what a citation file should look like and what it should be called.
\item Test adoption, for example, with LBNL.
\item Create credit file for a set of software.
\item Build a validator (and perhaps a visualizer) for credit files.
\item Write a tool to collect files and visualize/output interconnections (which 
software is used with which), based on an existing project.
\item When we (the group members) write papers, we should track the software 
we use, and encourage the software developers to make their software citable 
and create credit files.
\item Build a tool to export the credit file to BibTex and other citation styles.
\item Make sure the BibTex entries (exported from internal data) are somewhat 
standardized so that they can be imported into papers.  Also make sure that 
standard LaTeX style files understand and accept these entries.
\end{enumerate}
